---
title: "HW6"
author: "Talea Cornelius"
date: "11/19/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RCurl)
library(purrr)
library(modelr)
library(mgcv)
```


> Problem 1

> Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r load data}
homicide_dat <- read.csv(text = getURL("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")) 
names(homicide_dat)
```

```{r tidy data and select for city and race}
homicide_dat = homicide_dat %>%
  janitor::clean_names() %>%
  filter(victim_race == "Black" | victim_race == "White") %>%
  mutate(state = ifelse(city == "Tulsa", "OK", state)) %>%
  mutate(city_state = paste(city, sep = ", ", str_to_upper(state)),
         resolved = ifelse(disposition == "Closed by arrest", 1, 0),
         victim_age = as.numeric(victim_age),
         victim_race = factor(victim_race, levels = c("White", "Black"))) %>%
  group_by(city_state) %>%
  select(city_state, everything())
```

> For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r logistic regression for baltimore}

baltimore = homicide_dat %>% 
  subset(city_state == "Baltimore, MD") %>%
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., 
      family = binomial()) 

baltimore %>%
  broom::tidy() %>%
  filter(term == "victim_raceBlack") %>%
  mutate(term = str_replace(term, "victim_raceBlack", "Black v. White Race"),
         OR = exp(estimate), 
         lower_CI = exp(estimate - (1.96*std.error)),
         upper_CI = exp(estimate + (1.96*std.error))) %>%
  select(term, OR, lower_CI, upper_CI)
         

```

> Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing Black victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r logistic regression for each city}
logistic_df = homicide_dat %>%
  nest(data = -city_state) %>%
  mutate(models = map(data, ~glm(resolved ~ victim_age + victim_race + victim_sex,
                                 data = ., family = binomial())),
         models = map(models, broom::tidy)) %>% 
  select(-data) %>% 
  unnest(models) %>%
  filter(term == "victim_raceBlack") %>%
  mutate(term = str_replace(term, "victim_raceBlack", "Black v. White Race"),
         OR = exp(estimate), 
         lower_CI = exp(estimate - (1.96*std.error)),
         upper_CI = exp(estimate + (1.96*std.error))) %>%
  select(term, OR, lower_CI, upper_CI) 

```

> Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r plot ORs and CIs}
logistic_df %>% 
  ggplot(aes(x = reorder(city_state, OR), y = OR, color = OR)) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower_CI, 
                    ymax = upper_CI)) +
  labs(
    title = "Association of Black v. White Race with Resolved Status of Homicide",
    x = "City",
    y = "Odds Ratio and 95% CI"
  ) +
  theme(legend.position="none") +
  theme(text = element_text(size=10),
        axis.text.x = element_text(angle = 90, hjust=1, size=8)) +
  scale_color_gradient(low="purple", high="orange")

```


> Problem 2

> In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

>babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)

> Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r load and tidy birthweight data}
birthweight_dat = read_csv(file = "./data/birthweight.csv",
                       col_types = cols(
  .default = col_double(),
  `babysex` = col_factor(),
  `frace` = col_factor(),
  `mrace` = col_factor(),
  `malform` = col_factor())) %>%
  janitor::clean_names() %>%
  mutate(frace = recode(frace, "1" = "White", "2" = "Black", 
           "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
         mrace = recode(mrace, "1" = "White", "2" = "Black", 
           "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
         babysex = recode(babysex, "1" = "Male", "2" = "Female"),
         malform = recode(malform, "1" = "Present", "0" = "Absent")) %>%
  select(-c(pnumlbw, pnumsga)) #removed due to zero variability

sum(is.na(birthweight_dat)) #no missing data

```

> Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r select candidate predictors}
birthweight_dat %>%
  lm(bwt ~ ., data = .) %>%
  broom::tidy() %>%
  select(term, estimate, p.value) %>%
  filter(p.value <= .1)

```

**Based on prior research, I anticipate that gestational weeks, maternal race, and maternal smoking are each associated with birthweight. I additionally conducted a regression model with all predictors, and included those that were found to be significant at p < .10, two-tailed. Thus, my final predictor variables included: (1) baby sex, (2) baby head circumference, (3) baby length, (4) maternal weight at delivery, and (5) parity, as well as (6) gestational weeks, (7) maternal race, and (8) maternal smoking.**

```{r selected model}

birthweight_dat %>%
  lm(bwt ~ babysex + bhead + blength + delwt + 
       gaweeks + mrace + parity + smoken, data = .) %>%
  broom::tidy()

```

> Compare your model to two others:

> One using length at birth and gestational age as predictors (main effects only), and one using head circumference, length, sex, and all interactions (including the three-way interaction) between these. Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r create cross validation dataset}
cv_df =
  crossv_mc(birthweight_dat, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

```{r gather prediction errors}
cv_df = 
  cv_df %>% 
  mutate(
    my_mod  = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + 
       gaweeks + mrace + parity + smoken, data = .x)),
    mod_v1  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    mod_v2  = map(train, ~lm(bwt ~ bhead*blength*babysex, data = .x))) %>% 
  mutate(
    rmse_my_mod = map2_dbl(my_mod, test, ~rmse(model = .x, data = .y)),
    rmse_mod_v1 = map2_dbl(mod_v1, test, ~rmse(model = .x, data = .y)),
    rmse_mod_v2 = map2_dbl(mod_v2, test, ~rmse(model = .x, data = .y)))
```

```{r plot prediction error}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, color = "red")) + geom_violin()
```

**From this plot, it is apparent that the model including only length at birth and gestational age is the worst fitting model. My model has the lowest cross-validated prediction error, and is therefor the preferred model, however, the model including head circumference, length, sex, and all two- and three-way interactions also performs decently well.**


> Problem 3

> For this problem, we’ll use the 2017 Central Park weather data that we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

r^2
log(β^0∗β^1)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1). Note: broom::glance() is helpful for extracting r^2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β^0∗β^1).